---
title: "Untitled"
author: "datadavidz"
date: "4/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(lubridate)
library(GGally)
library(factoextra)

here::i_am("moneyball_part1.Rmd")
```

Based on the TidyX screencasts, for classifying pitches based on pitch f/x data.

Load the rds files containing the original data
```{r}
set1 <- readRDS(file = here("data", "2016_04_21_to_2016_04_23_pitch.rds"))
set2 <- readRDS(file = here("data", "2016_04_24_to_2016_04_26_pitch.rds"))

train <- bind_rows(set1, set2)
rm(set1, set2)

test <- readRDS(file = here("data", "2016_04_27_to_2016_04_30_pitch.rds"))

names(train)
```
*Potential Variables to retain in the analysis*

* start_speed (speed at 0 - 50ft)
* end_speed (speed at the front of home plate)
* pfx_x (a measure of break of the pitch in the x-direction)
* pfx_z (a measure of break of the pitch in the y-direction)
* px (a measure of x-value of position at the front of home plate)
* pz (a measure of the z-value of position at the front of home plate)
* x0 (x position at 50ft)
* z0 (z position at 50ft)
* vx0 (x velocity at 50ft)
* vz0 (z velocity at 50ft)

```{r}
skimr::skim(train)
```
```{r}
train %>%
  count(pitch_type, sort = TRUE)
```

**How many pitch types are there?**

* Remove NA
* Remove KN, IN, PO, EP, and FO

```{r}
train_cleaned <- train %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO"),
         !is.na(pitch_type)) %>%
  select(pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0) %>%
  drop_na()

test_cleaned <- test %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO", "SC"),
         !is.na(pitch_type)) %>%
  select(pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0) %>%
  drop_na()

skimr::skim(train_cleaned)
```
*Pitch Types*

* FF = four-seam fastball
* SL = slider
* CH = change up
* FT = two-seam fastball
* SI = sinker
* CU = curveball
* FC = cutter
* KC = knuckle-curve
* FS = split-finger fastball

```{r}
pitch_boxplot_func <- function(x_var){
  train_cleaned %>%
    ggplot(aes(x = x_var, y = reorder(pitch_type, x_var))) +
    geom_boxplot()
}
pitch_boxplot_func(x_var = train_cleaned$start_speed)
pitch_boxplot_func(x_var = train_cleaned$end_speed)
pitch_boxplot_func(x_var = train_cleaned$pfx_x)
pitch_boxplot_func(x_var = train_cleaned$pfx_z)
pitch_boxplot_func(x_var = train_cleaned$px)
pitch_boxplot_func(x_var = train_cleaned$pz)
pitch_boxplot_func(x_var = train_cleaned$x0)
pitch_boxplot_func(x_var = train_cleaned$z0)
pitch_boxplot_func(x_var = train_cleaned$vx0)
pitch_boxplot_func(x_var = train_cleaned$vz0)
```



Could possibly carry forward the date
```{r}
# train %>% select(y) %>% mutate(y2 = str_remove(y, "T.*$"),
#                                y2 = ymd(y2))
```

```{r}
ggcorr(train_cleaned, label = TRUE, size = 3)
```
Next we look at summarizing
```{r}
pitch_type_summary <- train_cleaned %>%
  group_by(pitch_type) %>%
  summarize(across(everything(), .fns = median))

pitch_type_summary
```

```{r}
df <- pitch_type_summary %>% select(-pitch_type) %>% as.data.frame()

# Set rowname: 
row.names(df) <- pitch_type_summary %>% pull(pitch_type)

# Compute distances: 
dd <- dist(df, method = "euclidean")

# Visualize the dissimilarity: 
fviz_dist(dd, lab_size = 12)
```

```{r}
hc <- hclust(dd, method = "ward.D2")

# Create a draft of dendrogram by using fviz_dend() function: 

fviz_dend(hc, 
          k = 4,   
          cex = 0.5, 
          rect = TRUE, 
          rect_fill = TRUE, 
          horiz = FALSE, 
          palette = "jco", 
          rect_border = "jco", 
          color_labels_by_k = TRUE)
```

```{r}
# Cut tree into 4 groups: 
sub_grp <- cutree(hc, k = 4)

# Create plot of clusters: 
fviz_cluster(list(data = df, cluster = paste0("Group", sub_grp)), 
             alpha = 1, 
             palette = "jco", 
             labelsize = 9, 
             ellipse.type = "norm")
```

Make a summary table
```{r}
bind_cols(pitch_type_summary, clust_num = sub_grp) %>%
  group_by(clust_num) %>%
  summarize(across(.cols = start_speed:vz0, .fns = list(avg = mean, sd = sd)), .groups = "drop") %>%
  pivot_longer(cols = -clust_num) %>%
  pivot_wider(names_from = clust_num, values_from = value)
```
Last you can do hclust on all the training data
```{r}
clust_df <- train_cleaned %>% select(-pitch_type) %>% as.data.frame()

# Set rowname: 
#row.names(clust_df) <- train_cleaned %>% pull(pitch_type)

# Compute distances: 
pitch_clust <- dist(clust_df, method = "euclidean")

## select a disimilarity metric
pitch_clust_final <- hclust(pitch_clust, method = "ward.D2")

## get the cluster assignments
clust_assignment <- cutree(pitch_clust_final, k = 4)

## put the results in a data frame
pitch_clust_df <- bind_cols(train_cleaned,
                            cluster_num = clust_assignment)

## summarize clusters
pitch_clust_df %>%
  select(-pitch_type) %>%
  group_by(cluster_num) %>%
  summarize(across(.cols = start_speed:vz0,
                  list(avg = mean, sd = sd))) %>%
  pivot_longer(-cluster_num) %>%
  pivot_wider(names_from = cluster_num,
              values_from = value)

## Which pitches in each cluster
table("Pitch Type" = pitch_clust_df$pitch_type, "Cluster Number" = pitch_clust_df$cluster_num)
```

```{r}
pitch_clust_df %>%
  count(cluster_num, pitch_type) %>%
  mutate(group = case_when(pitch_type %in% c("FF", "FT", "FC", "SI", "FS") ~ "Fastball Pitches",
                           TRUE ~ as.character("Off-Speed Pitches"))) %>%
  ggplot(aes(x = as.factor(cluster_num), y = n, fill = group)) +
  geom_col()
```
