---
title: "Untitled"
author: "datadavidz"
date: "5/11/2021"
output: html_document
---

My own version of TidyX Episode 55 where we fit decision tree and random forest models to the pitch f/x data.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)
library(here)
library(lubridate)

library(rpart.plot)

here::i_am("moneyball_part2.Rmd")
```

Load data, minor preprocessing, and do the train/test split using rsample methodology.
Calculation of the split proportion by date/time could be simplified to just taking 2/3 but wanted to be consistent with TidyX.
```{r}
#training
set1 <- readRDS(file = here("data", "2016_04_21_to_2016_04_23_pitch.rds"))
set2 <- readRDS(file = here("data", "2016_04_24_to_2016_04_26_pitch.rds"))
#testing
set3 <- readRDS(file = here("data", "2016_04_27_to_2016_04_30_pitch.rds"))

#Keeping date and time
pitch_raw <- set1 %>%
  bind_rows(set2) %>%
  bind_rows(set3) %>%
  mutate(pitch_date = str_replace(y, "T", " "),
         pitch_date = str_remove(pitch_date, "\\..*$"),
         pitch_date = ymd_hms(pitch_date))

#Need to know the max date for set 2 for splitting purpose
set2_max_date <- set2 %>%
  mutate(pitch_date = str_replace(y, "T", " "),
         pitch_date = str_remove(pitch_date, "\\..*$"),
         pitch_date = ymd_hms(pitch_date)) %>%
  summarize(max_date = max(pitch_date, na.rm = TRUE)) %>%
  pluck(., 1)

pitch_clean <- pitch_raw %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO", "SC"),
         !is.na(pitch_type)) %>%
  select(pitch_date, pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0) %>%
  drop_na()

prop <- nrow(pitch_clean %>% filter(pitch_date <= set2_max_date )) / nrow(pitch_clean)
```

Create the initial split based on time (like done for TidyX)
```{r}
pitch_split <- initial_time_split(data = pitch_clean, prop = prop)
pitch_train <- training(pitch_split)
pitch_test <- testing(pitch_split)
```

Create the recipe
```{r}
rpart_rec <- recipe(pitch_type ~ ., data = pitch_train) %>%
  update_role(pitch_date, new_role = "id") %>%
  prep()

rpart_rec
```
Specify the decision tree model
```{r}
rpart_spec <- decision_tree() %>%
   set_engine("rpart") %>%
   set_mode("classification")

rpart_spec
```

Create the workflow
```{r}
rpart_wf <- workflow() %>%
  add_recipe(rpart_rec) %>%
  add_model(rpart_spec)
```

Fit the decision tree
```{r}
rpart_res <- fit(rpart_wf, data = pitch_train)

rpart_res
```

Plot the decision tree
```{r}
rpart.plot(rpart_res$fit$fit$fit, roundint = FALSE) #yep - not ideal
```

Do the fit and the predictions based on the initial data split.
```{r}
rpart_fit <- rpart_wf %>% last_fit(pitch_split, metrics = metric_set(accuracy, kap))

rpart_fit %>% collect_metrics()
```
Show the confusion matrix for pitch predictions
```{r}
rpart_fit %>% collect_predictions() %>% conf_mat(truth = pitch_type, estimate = .pred_class)
```

### Random Forest Model













