---
title: "Untitled"
author: "datadavidz"
date: "4/30/2021"
output: html_document
---

My own version of TidyX Episode 54 where we do KNN and UMAP clustering of Pitch F/X data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)
library(here)
library(lubridate)

library(embed) #for UMAP

here::i_am("moneyball_part2.Rmd")
```

Load data, minor preprocessing, and do the train/test split using rsample methodology.
Calculation of the split proportion by date/time could be simplified to just taking 2/3 but wanted to be consistent with TidyX.
```{r}
#training
set1 <- readRDS(file = here("data", "2016_04_21_to_2016_04_23_pitch.rds"))
set2 <- readRDS(file = here("data", "2016_04_24_to_2016_04_26_pitch.rds"))
#testing
set3 <- readRDS(file = here("data", "2016_04_27_to_2016_04_30_pitch.rds"))

#Keeping date and time
pitch_raw <- set1 %>%
  bind_rows(set2) %>%
  bind_rows(set3) %>%
  mutate(pitch_date = str_replace(y, "T", " "),
         pitch_date = str_remove(pitch_date, "\\..*$"),
         pitch_date = ymd_hms(pitch_date))

#Need to know the max date for set 2 for splitting purpose
set2_max_date <- set2 %>%
  mutate(pitch_date = str_replace(y, "T", " "),
         pitch_date = str_remove(pitch_date, "\\..*$"),
         pitch_date = ymd_hms(pitch_date)) %>%
  summarize(max_date = max(pitch_date, na.rm = TRUE)) %>%
  pluck(., 1)

pitch_clean <- pitch_raw %>%
  filter(!pitch_type %in% c("KN", "IN", "PO", "EP", "FO", "SC"),
         !is.na(pitch_type)) %>%
  select(pitch_date, pitch_type, start_speed, end_speed, pfx_x, pfx_z, px, pz, x0, z0, vx0, vz0) %>%
  drop_na()

prop <- nrow(pitch_clean %>% filter(pitch_date <= set2_max_date )) / nrow(pitch_clean)
```

Create the initial split based on time (like done for TidyX)
```{r}
pitch_split <- initial_time_split(data = pitch_clean, prop = prop)
pitch_train <- training(pitch_split)
pitch_test <- testing(pitch_split)
```

Create the recipe
```{r}
pitch_rec <- recipe(pitch_type ~ ., data = pitch_train) %>%
  update_role(pitch_date, new_role = "id") %>%
  step_normalize(all_numeric()) %>%
  prep()

pitch_rec
```

Create cross-validation folds for tuning
```{r}
set.seed(234)

pitch_folds <- vfold_cv(pitch_train, v = 10)

pitch_folds
```


Create the model specification
```{r}
knn_spec <- nearest_neighbor(neighbors = tune()) %>%
  set_engine("kknn") %>%
  set_mode("classification")
```

Create the tune grid
```{r}
knn_grid <- grid_regular(neighbors(), levels = 10)

knn_grid
```

Create the workflow
```{r}
knn_wf <- workflow() %>%
  add_recipe(pitch_rec) %>%
  add_model(knn_spec)
```

Tune the nearest neighbors using folds
```{r}
doParallel::registerDoParallel()

knn_tune <- tune_grid(
  knn_wf,
  resamples = pitch_folds,
  grid = knn_grid,
  metrics = metric_set(accuracy, kap)
)
```

Analyze tune results
```{r}
knn_tune %>% show_best("kap", n = 5)
knn_tune %>% show_best("accuracy", n = 5)

knn_tune %>% autoplot()
```

Finalize neighbors parameter at 8.  This could also be done by select_best but to be consistent with TidyX we use 8.
```{r}
knn_final_wf <- knn_wf %>% finalize_workflow(tibble(neighbors = 8))
```

```{r}
knn_final_fit <- knn_final_wf %>% last_fit(pitch_split, metrics = metric_set(accuracy, kap))

knn_final_fit %>% collect_metrics()
```

Show the confusion matrix for pitch predictions
```{r}
knn_final_fit %>% collect_predictions() %>% conf_mat(truth = pitch_type, estimate = .pred_class)

```
```{r}
knn_final_fit %>% collect_predictions() %>% conf_mat(truth = pitch_type, estimate = .pred_class) %>% summary()
```


Misclassification rate by pitch
```{r}
knn_final_fit %>% 
  collect_predictions() %>% 
  select(observed = pitch_type, predicted = .pred_class) %>%
  group_by(observed, predicted) %>%
  summarize(freq = n(), .groups = "drop") %>%
  group_by(observed) %>%
  summarize(mcr = 1- freq[observed == predicted] / sum(freq), 
            n = sum(freq), .groups = "drop")
```

### UMAP

UMAP can be performed as a preprocessing step (recipe) in tidymodels framework using the ```embed``` library.

```{r}
umap_rec <- recipe(~ ., data = pitch_train) %>% #no outcome - unsupervised learning
  update_role(pitch_type, pitch_date, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_umap(all_predictors()) %>%
  prep()
```

```{r}
pitch_conv <- tribble(~pitch_type, ~pitch_label,
                      "CH", "Change-up",
                      "CU", "Curveball",
                      "FC", "Cutter",
                      "FF", "Four-seam Fastball",
                      "FS", "Split-finger Fastball",
                      "FT", "Two-seam Fastball",
                      "KC", "Knuckle Curve",
                      "SI", "Sinker",
                      "SL", "Slider")
```

Create a plot to analyze the umap projection
```{r}
juice(umap_rec) %>%
  left_join(pitch_conv, by = "pitch_type") %>%
  ggplot(aes(umap_1, umap_2)) +
  geom_point(aes(color = pitch_label), alpha = 0.7, size = 2) +
  labs(color = NULL)
```

